Care Phenotypes: A Novel Approach to Understanding Healthcare Data Collection Patterns

### 2025-03-29 docs/manuscript/manuscript.tex
V2.

\section{Methods}
% Methods content goes here

>>>>

\section{Methods}

\subsection{Data Processing Framework}

We developed a comprehensive framework for processing MIMIC-IV data, implemented as a Python package. The framework consists of several key components:

\subsubsection{Data Structures and Formats}

We defined standardized data structures for various MIMIC data types, including:
\begin{itemize}
    \item Patient demographics and admission information
    \item Laboratory measurements and chart events
    \item ICU stays and clinical scores
\end{itemize}

These structures ensure type safety and consistency throughout the data processing pipeline. We implemented robust data validation and integrity checks to maintain data quality.

\subsubsection{Clinical Score Calculations}

Our framework includes implementations of several widely-used clinical scoring systems:

\begin{itemize}
    \item \textbf{SOFA (Sequential Organ Failure Assessment)}: Evaluates organ dysfunction across six systems (respiratory, coagulation, liver, cardiovascular, CNS, and renal)
    \item \textbf{Charlson Comorbidity Index}: Assesses patient comorbidity burden using 17 weighted conditions
    \item \textbf{APACHE II}: Comprehensive scoring system incorporating acute physiology, chronic health, and age components
    \item \textbf{SAPS II}: Simplified acute physiology scoring system
    \item \textbf{Elixhauser Comorbidity Index}: Detailed assessment of 31 comorbidities
\end{itemize}

Each scoring system is implemented as a modular component, allowing for flexible integration and extension. The implementations handle missing data gracefully and provide detailed component-level analysis.

\subsubsection{Data Processing Pipeline}

The data processing pipeline includes:
\begin{itemize}
    \item Standardized data loading and validation
    \item Automated data cleaning and normalization
    \item Efficient handling of time-series data
    \item Integration of multiple data sources
    \item Comprehensive error handling and logging
\end{itemize}

\subsection{Implementation Details}

Our implementation focuses on:
\begin{itemize}
    \item \textbf{Modularity}: Each component is self-contained and follows consistent interfaces
    \item \textbf{Type Safety}: Comprehensive type hints and validation
    \item \textbf{Documentation}: Detailed docstrings and usage examples
    \item \textbf{Performance}: Optimized data structures and algorithms
    \item \textbf{Extensibility}: Easy addition of new scoring systems and data types
\end{itemize}
